---
main_topsize: 0.2 #percent coverage of the poster
main_bottomsize: 0.1
#ESSENTIALS
title: "How does development aid impact Nigeria's maternal mortality rate?"
author:
  - name: 'Camryn Allen'
    affil: 1
    main: true
    orcid: '0000-0002-1099-3857'
    email: ca9134a@american.edu
affiliation:
  - num: 1
    address: Department of Politics, Governance, and Economics | School of International Service (SIS) | American University
main_findings:
  - "Nigeria's Astronomical Maternal Deaths Unaffected by World Bank Aid"
primary_colour: "#925653"
secondary_colour: "#F7A5A5"
accent_colour: "#841921"
output: 
  posterdown::posterdown_betterport:
    self_contained: yes
    pandoc_args: --mathjax
    number_sections: false
bibliography: packages.bib
link-citations: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html") 

library(dplyr)
library(ggplot2)
library(readr)
library(tidyverse)
library(rmdformats)
library(patchwork)
library(reshape2)
```

# Introduction

The World Health Organization defines a "maternal death" as "a female death from any cause related to or aggravated by pregnancy or its management (excluding accidental or incidental causes) during pregnancy and childbirth or within 42 days to a year after termination of pregnancy." This project aims to examine how annual maternal deaths in Nigeria are impacted by World Bank (WB) aid to the country. Exploring the effect of foreign aid from leading international governance institutions, such as the World Bank, on maternal deaths in the most devastatingly affected areas, such as sub-Saharan Africa, is a crucial necessity if maternal deaths are to be effectively addressed and resolved everywhere. The expectation is that the World Bank's annual aid to Nigeria will have a negative impact on the country's maternal deaths (i.e. an increase in WB aid will lead to a decrease in Nigeria's maternal deaths).

```{r, include=FALSE}
knitr::write_bib(c('posterdown', 'rmarkdown','pagedown'), 'packages.bib')
```

# Methods

Data reflecting Nigeria's annual maternal deaths between the years 2000 - 2020 and World Bank aid (USD) to Nigeria within the same time frame have been collected from the World Health Organization (2023) "Number of Maternal Deaths - Nigeria" data set and the World Bank's "Net Official Development Assistance and Official Aid Received (Current US$) - Nigeria" data set, respectively. 

Key variables include:

```{r, echo=TRUE, eval=FALSE}
Deaths: annual maternal deaths within Nigeria
AidDollars: WB annual aid to Nigeria
Year: Individual year of data collection
```

```{r, eval=TRUE, echo=FALSE, fig.height=6, fig.cap='This bar graph depicts data of 20 African countries with the highest maternal deaths in 2019 from the World Health Organization ["Number of Maternal Deaths - Nigeria" data set](https://data.worldbank.org/indicator/SH.MMR.DTHS?locations=NG) (2023).'}
#MAIN GRAPH: MATERNAL DEATHS ACROSS COUNTRIES
MatDeath = read_csv('MatDeath.csv')

AfrMatDeath = 
  MatDeath %>%
  filter(`Country Name` %in% c('Angola', 'Benin', 'Bostswana', "Cote d'Ivoire", 'Cameroon', 'Congo, Dem. Rep.', 'Ethiopia', 'Ghana', 'Guinea', 'Equitorial Guinea', 'Kenya', 'Madagascar', 'Mali', 'Malawi', 'Niger', 'Nigeria', 'Sudan', 'Somalia', 'South Sudan', 'Chad', 'Tanzania', 'Uganda'))

AfrMatDeath$`Indicator Name` <- NULL
AfrMatDeath$`Country Code` <- NULL
AfrMatDeath$`Indicator Code` <- NULL
names(AfrMatDeath)[names(AfrMatDeath) == "2020"] <- "CovidYear"
names(AfrMatDeath)[names(AfrMatDeath) == "2019"] <- "CovidStart"
AfrMatDeath$`Country Name` <- reorder(AfrMatDeath$`Country Name`, +AfrMatDeath$CovidStart)

AfrMatDeath |>
  ggplot(aes(x = CovidStart, y = `Country Name`))+ 
  geom_bar(stat = 'identity', fill = '#E4257A', width = 0.85) + 
  ggtitle('Top 20 Leading Maternal Deaths in Africa (2019)') + 
  labs(
    x = 'Annual Maternal Deaths',
    y = 'Country'
  ) + 
  geom_text(aes(label = CovidStart), vjust = 0.5, hjust = -0.1, size = 2.75) +
  scale_x_continuous(
    expand = expansion(mult = c(0, 0.08)),
    breaks = c(0, 20000, 40000, 60000, 80000),
    labels = c('0', '20,000', '40,000', '60,000', '80,000')
  ) +
  theme_bw() + 
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(
      hjust = 0.5,
      face = 'bold'),
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(),
    axis.title.x = element_text(face = 'bold'),
    axis.title.y = element_text(face = 'bold')
  ) 
```



```{r, echo=FALSE, eval=TRUE, fig.cap='Nigeria maternal deaths between 2000 - 2020 have been isolated from the World Bank [Net Official Development Assistance and Official Aid Received (Current US$) - Nigeria data set](https://data.worldbank.org/indicator/DT.ODA.ALLD.CD?end=2020&locations=NG&start=2000).'}
library(readxl)
library(readr)
NMDeaths <- read_excel('NMDeaths.xlsx')
WBNAid <- read_excel('WBNAid.xlsx')

DeathAid <- merge(NMDeaths, WBNAid, by = 'Year', all = TRUE)
names(DeathAid)[names(DeathAid) == "WB Aid"] <- "AidDollars"

NMDPlot = 
  NMDeaths |>
  ggplot(aes(x = Year, y = Deaths)) + 
  geom_line(
    color = '#E4257A'
  ) +
  geom_point() + 
  ggtitle('Maternal Deaths in Nigeria (2000 - 2020)') + 
  labs(
    y = 'Annual Maternal Deaths'
  ) +
  scale_y_continuous(
    breaks = c(60000, 65000, 70000, 75000, 80000, 85000), #How to make y axis start at 60K?
    labels = c('60K', '65K','70K', '75K', '80K', '85K')
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = 'bold'),
    axis.title.x = element_text(face = 'bold'),
    axis.title.y = element_text(face = 'bold')
  )
NMDPlot #How to label points at 2000, 2005, 2010, 2015, 2020?

```



# Results

```{r MinMaxTable, echo=FALSE, eval=TRUE}
library(readxl)
library(readr)
library(knitr)
library(kableExtra)
NMDeaths <- read_excel('NMDeaths.xlsx')
WBNAid <- read_excel('WBNAid.xlsx')
DeathAid <- merge(NMDeaths, WBNAid, by = 'Year', all = TRUE)
names(DeathAid)[names(DeathAid) == "WB Aid"] <- "AidDollars"

DeathAidSum <- data.frame( #USE THIS TABLE!
  Minimum = sapply(DeathAid, min),
  Maximum = sapply(DeathAid, max)
)

kable(DeathAidSum)
```
The minimum maternal death count in Nigeria between the years 2000-2020 was 61,000, while the maximum reached 86,000 in 2019 (presumably affected by the COVID-19 pandemic). Additionally, the minimum World Bank annual aid to Nigeria was more than 167M USD in 2001 and maxes at more than 11.4B USD in 2006.  

```{r, echo=FALSE, eval=TRUE}
library(readxl)
library(readr)
NMDeaths <- read_excel('NMDeaths.xlsx')
WBNAid <- read_excel('WBNAid.xlsx')

DeathAid <- merge(NMDeaths, WBNAid, by = 'Year', all = TRUE)
names(DeathAid)[names(DeathAid) == "WB Aid"] <- "AidDollars"
DeathAid <- DeathAid[DeathAid$AidDollars < 6000000000, ] #How to fix the Y axis labels?? Do I limit or have entire dataset?
DeathAid$Money <- DeathAid$AidDollars / 1000000000

ggplot(DeathAid, aes(y = Deaths, x = Money)) +
  geom_point() +
  geom_smooth(method = 'lm', se = F, color = '#E4257A') +
  ggtitle('A Relationship between Maternal Deaths and World Bank Aid in Nigeria') + 
  labs(
    x = 'Aid USD (Billions)',
    y = 'Annual Maternal Deaths'
  ) +
  scale_y_continuous(
    breaks = c(60000, 65000, 70000, 75000, 80000, 85000), #How to make y axis start at 60K?
    labels = c('60K', '65K','70K', '75K', '80K', '85K')
  ) +
  theme_bw() + 
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = 'bold'),
    axis.title.x = element_text(face = 'bold'),
    axis.title.y = element_text(face = 'bold')
  )

```

The graph above reveals a positive correlation between WB aid and maternal deaths in Nigeria throughout 2000 - 2020. Despite the statistical tests reflecting that WB aid has an extremely minimal influence on maternal deaths in Nigeria, it is important to emphasize the trend of worsening humanitarian issues attracting increasingly higher amounts of foreign aid in hopes of aggressively seeking solutions.  Visible in the graph above, Nigeria's maternal deaths are seemingly unaffected by increasing influxes of development aid.From these results, it can be concluded that WB aid has designated and raised development aid to Nigeria in general over the past two decades, however this aid has not yet caused a significant decrease in Nigeria's annual maternal deaths. More research should be conducted to explore root causes and obstacles that explain why gradual increases in development aid to Nigeria correlate to or align with a continual increase in maternal death in Nigeria.



# References

The World Bank (2024). "Net Official Development Assistance and Official Aid Received (current US$) - Nigeria", Nigeria: https://data.worldbank.org/indicator/DT.ODA.ALLD.CD?end=2020&locations=NG&start=2000 

The World Health Organizaztion (2023). "Number of Maternal Deaths - Nigeria", Nigeria, The World Bank: https://data.worldbank.org/indicator/SH.MMR.DTHS?locations=NG 